Tunable VIMU frame from averaging multiple Non-collocated IMUs

### 2024-09-09 paper/main.tex
Added example trajectory image for real-world experiment.
Finished draft and proof read.

\title{Weighted Averaging of Multiple Non-collocated Inertial Measurement Units}

@@

\begin{abstract}
Multi-IMU fusion has been an effective way of improving MEMS IMU accuracy. In this article, we discuss the averaging technique for multiple IMUs with significant separation. We also present methods of selecting weighting terms such that we either minimize the uncertainty in fused sensor output or arbitrarily placing the reference frame of the combined virtual IMU (VIMU). We tested our method with simulation and evaluated on real dataset. The result shows that the averaging technique works for IMUs with separation and performance gain is observed.
\end{abstract}

@@

Inertial measurement unit combining a gyroscope and accelerometer that provides measurement of angular velocity and linear acceleration has wide application in robotics and UAS. By integrating the measurement, one can track the relative orientation and position of the system to certain degree of accuracy within reasonable time frame \cite{tang2022_preintegration}. However due to the interoceptive nature of IMU, it is often paired with additional sensors like camera, lidar, or GNSS to reduce drift \cite{alaba2024_gps_and_imu, shan2020_liosam, he2020_visual_imu_and_gps}. In general there are two main approaches to the tracking problem with an IMU - filter-based approach vs. graph optimization approach.

In filter-based approach, IMU measurement is used as process input to an Extended Kalman Filter in propagation step. Then the measurement from other exteroceptive sensors are used for state correction in update step. One such example is MSCKF (Multi-State Constraint Kalman Filter) proposed in \cite{Anastasios2007_MSCKF} which uses IMU for propagation and camera for correction. It proposed a measurement model that directly relate two robot poses based on the landmark observation in camera but independent from the landmark position itself. This allows for efficient tracking without including position of landmarks as part of the state vector as in traditional EKF-SLAM.

For graph optimization approach, such as VINS-MONO \cite{qin2018_vins-mono}, a pose graph is constructed by solving tightly-coupled, nonlinear optimization problem with preintegrated IMU measurement and camera observations. Pose graph optimization is run at the end to ensure global consistency.

Among many sensor configuration with IMU, monocular camera plus IMU is one of the minimal configuration that is well-studied \cite{qin2018_vins-mono, 10616216}. Also IMU has been used heavily as a complimentary sensor for GPS in GNSS-denied environment \cite{1008998, 8987949}. Based on these minimal configurations, various extension has been made. UMS-VINS \cite{jiang2023_UMS-VINS} provides a unified framework for using monocular and stereo cameras in visual-inertial odometry. Also the MSCKF framework \cite{Anastasios2007_MSCKF} also works for multiple cameras with a IMU. There are also works that aim to provide a complete framework to incorporate multiple different sensors in one unified system \cite{10587194}.

Despite works mentioned above, most system and application still only use single IMU in the estimation pipeline. To improve robustness, there have being works in introducing redundant IMUs for fault detection to build fault tolerance system. To detect a measurement error, \cite{Sturza1988_redundant} proposed parity vector based on hypothesis testing. Optimal geometric configuration of multiple redundant IMUs is also explored in \cite{Colomina2004REDUNDANTIF, guerrier2009, xue2023}.

% Further work discuss how to improve the accuracy by fusing multiple low cost IMUs \cite{waegli2008, patel2022_multi-imu, xue2023}.

 In this article the focus is on improving the accuracy of the overall system by leveraging the multiple redundant IMUs onboard in accordance with other sensors provided.

 % As summarized in \cite{patel2022_multi-imu}, for closely place IMUs it is convenient to simply average all IMU measurements after aligning them under a common reference frame.

@@

In this section, we will introduce some prior works on improving the accuracy of multi-IMU system.

@@

One category of approach is to perform fusion of multiple IMU before using the measurement in inertial navigation system (INS). For example, \cite{jafari2014_PEM} uses Prediction Error Minimization method to model noise of IMUs and use Kalman filter to estimate the optimal process signal. \cite{xue2023} proposed to use Kalman filter to estimate the angular velocity measured by multiple non-orthogonal gyroscopes. Instead of using Kalman filter, an easier approach would be to directly average multiple gyroscopes after aligned under a common virtual IMU frame as discussed in \cite{waegli2008, patel2022_multi-imu, Colomina2004REDUNDANTIF}.

However, averaging of multiple IMUs was not considered due to the additional acceleration terms from rotation (lever arm effect), unless they are closed place and the separation in between is ignored.

@@

Typically state estimation problem with single IMU involves tracking the orientation, linear velocity and linear acceleration of the vehicle (commonly defined at the IMU). With additional IMUs, additional state variables like linear acceleration, angular velocity and angular acceleration are added. This allows using the full mechanization model of accelerometer as measurement update in Kalman filter as in \cite{Bancroft2011DataFA, Beaudoin2018_satelite}.

@@

Another approach is to track individual IMU's state and associate them with geometric constraints \cite{waegli2008, Beaudoin2018_satelite}. Using multiple IMUs and cameras, \cite{Eckenhoff2021_MIMC-VINS} proposed to track all stacked states of base and auxillary IMUs using MSCKF \cite{Anastasios2007_MSCKF}. Assuming the calibrations are known, relative geometric constraints is imposed between states of different IMUs in the form of measurement update in addition to camera observations.

@@

As presented, multiple approaches have being proposed. \ref{constraint} would be computationally expensive if number of IMU are large as it has many states all in one Kalman filter. The averaging approach in \ref{VIMU} is the simplest to implement and doesn't significantly alter the existing single IMU framework. Although averaging between accelerometers that are physically separated is not visited due to the lever arm effect, in \ref{methodology} we present that it is still possible to average the accelerometers despite the separation.

@@

For any quantity $x$ We define $\bar{x} = x + b_x$ as the biased value and $\Tilde{x} = x + n_x$ as the noisy value. If it is both biased and noisy measurement, it is denoted as $y_x$. If a quantity is in the form $x(t)$ then it is a time-varying quantity, otherwise it is time-invariant.

For any point $\textbf{p}$ or vector we define as $\textbf{r}_i^{pi}$ in inertia frame $\mathcal{F}_i$. For example, the vehicle position in inertial frame thus become $\textbf{r}_i^{vi}$.

For rotation matrix or transformation matrix, we have $\textbf{C}_{vi}$ or $\textbf{T}_{vi}$ for transforming point $\textbf{r}_i^{pi}$ expressed in inertial frame $\bm{\mathcal{F}}_i$ into vehicle frame $\bm{\mathcal{F}}_v$ as $\textbf{r}_v^{pv}$.
$$
\left[\begin{matrix} \textbf{r}_v^{pv} \\ 1 \end{matrix}\right]
    = \textbf{T}_{vi} \left[\begin{matrix} \textbf{r}_i^{pi} \\ 1 \end{matrix}\right]
\quad \text{or} \quad
\textbf{r}_v^{pv} = \textbf{C}_{vi} \textbf{r}_i^{pi} + \textbf{r}_v^{iv}
$$

In the rest of the paper, we need to distinguish between body frame and vehicle frame. Body frame $\bm{\mathcal{F}}_b$ is for expressing sensor location relative to some mechanical datum and vehicle frame $\bm{\mathcal{F}}_v$ is for tracking vehicle pose in state estimation pipeline.

When referring the pose of a vehicle, it is a tuple of $\{\textbf{C}_{iv}, \textbf{r}_i^{vi}\}$ which forms the transformation matrix as follow,
$$
\textbf{T}_{iv} = \left[\begin{matrix}
    \textbf{C}_{iv} & \textbf{r}_i^{vi} \\
    \textbf{0} & 1
\end{matrix}\right]
$$

In this document $*^\wedge$ is specifically used for taking a $\mathbb{R}^3$ into skew symmetric matrix. While $*^\wedge$ is used for transforming an element of Lie algebra to corresponding element in Lie group and $*^\vee$ is used for the reverse operation.
$$
\bm{\omega}^\wedge = \left[\begin{matrix}
    0 & -\omega_2 & \omega_1 \\
    \omega_2 & 0 & -\omega_3 \\
    -\omega_1 & \omega_3 & 0
\end{matrix}\right]
$$

@@

In addition, $\textbf{g}$ is the acceleration from gravity and $\bm{\omega}$ and $\textbf{a}$ are the true angular velocity and true linear acceleration in vehicle frame. Thus the goal is to estimate $\bm{\omega}$ and $\textbf{a}$ from measurements of IMUs.

@@

Where $\textbf{b}_\omega$ is the bias of gyroscope measurement due to random walk and $\textbf{n}_\omega$ is the additive white noise of the measurement. Similarly we have the following measurement model for IMU at vehicle frame. Here $\textbf{b}_a$ is the bias of accelerometer and $\textbf{n}_a$ is the noise of the acceleration measurement.

@@

For IMU not at vehicle frame, measurement are in each IMU's local frame and we also have additional terms in acceleration measurement due to lever arm effect.
\begin{equation}
    \textbf{y}_{\omega,i}(t) = \textbf{C}_{i}^{-1} \bm{\omega}(t) + \textbf{b}_{\omega, i}(t) + \textbf{n}_{\omega,i}(t)
\end{equation}
\begin{equation}
\begin{split}
    \textbf{y}_{a,i}(t) &= \textbf{C}_i^{-1} \left[ \textbf{a}(t) + \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i) + \bm{\alpha}(t)^\wedge \textbf{r}_i \right] \\
    &+ \textbf{b}_{a,i}(t) + \textbf{n}_{a,i}(t)
\end{split}
\end{equation}

Here we have $\textbf{C}_i$ and $\textbf{r}_i$ short for $\textbf{C}_{vs}$ and $\textbf{r}_v^{sv}$ defining the pose $\textbf{T}_{vs}$ of the $i$-th IMU in vehicle frame (or VIMU frame as oppose to body frame).

@@

First we align gyroscope measurements with vehicle frame.
\begin{equation}
    \textbf{C}_{i} \textbf{y}_{\omega,i}(t) = \bm{\omega}(t) + \textbf{C}_{i} \textbf{b}_{\omega, i}(t) + \textbf{C}_{i} \textbf{n}_{\omega,i}(t)
\end{equation}

The weighted average ($\sum_i{a_i} = 1$) of all gyroscope is simply,
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_\omega(t) &= \sum_i{a_i \textbf{C}_{i} \textbf{y}_\omega(t)} \\
    &= \sum_i{a_i \left( \bm{\omega}(t) + \textbf{C}_{i} \textbf{b}_{\omega,i}(t) + \textbf{C}_{i} \textbf{n}_{\omega,i}(t) \right)} \\
    &= \bm{\omega}(t) + \sum_i{a_i \textbf{C}_{i} \textbf{b}_{\omega,i}(t)} + \sum_i{a_i \textbf{C}_{i} \textbf{n}_{\omega,i}(t)}
\end{split}
\end{equation}

@@

Align the all accelerometer measurements to vehicle frame.
\begin{equation}\label{eqn_accel}
\begin{split}
    \textbf{C}_{i} \textbf{y}_{a,i}(t) &= \textbf{a}(t) + \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i) + \bm{\alpha}(t)^\wedge \textbf{r}_i \\
    &+ \textbf{C}_{i} \textbf{b}_{a,i}(t) + \textbf{C}_{i} \textbf{n}_{a,i}(t)
\end{split}
\end{equation}

Summing equation (\ref{eqn_accel}) of all accelerometer we have,
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_a(t) &= \sum_i{a_i \textbf{C}_{i} \textbf{y}_a(t)} \\
    &= \textbf{a}(t) + \sum_i{a_i \textbf{C}_{i} \textbf{b}_{a,i}(t)} + \sum_i{a_i \textbf{C}_{i} \textbf{n}_{a,i}(t)} \\
    &+ \sum_i{a_i \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i)} + \sum_i{a_i \bm{\alpha}(t)^\wedge \textbf{r}_i} \\
\end{split}
\end{equation}

Using the associate rule $\textbf{a}^\wedge \textbf{b} + \textbf{a}^\wedge \textbf{c} = \textbf{a}^\wedge \left(\textbf{b} + \textbf{c}\right)$, we have,
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_a(t) &= \sum_i{a_i \textbf{C}_{i} \textbf{y}_a(t)} \\
    &= \textbf{a}(t) + \sum_i{a_i \textbf{C}_{i} \textbf{b}_{a,i}(t)} + \sum_i{a_i \textbf{C}_{i} \textbf{n}_{a,i}(t)} \\
    &+ \bm{\omega}(t)^\wedge \left(\bm{\omega}(t)^\wedge \left( \sum_i{a_i \textbf{r}_i} \right) \right) + \bm{\alpha}(t)^\wedge \left( \sum_i{a_i\textbf{r}_i} \right)\\
\end{split}
\end{equation}

If $\sum_i{a_i \textbf{r}_i} = \textbf{0}$ then we can eliminate the last two terms from lever arm effect. This gives,
\begin{equation}
    \bar{\textbf{y}}_a(t) = \textbf{a}(t) + \bar{\textbf{b}}_a(t) + \bar{\textbf{n}}_a(t)
\end{equation}

@@

As we can see, if we choose the vehicle frame and weights of accelerometers such that $\sum_i{a_i \textbf{r}_i} = \textbf{0}$, then we can ignore the separation of all IMUs and fused the accelerometer measurement through averaging.

\textbf{Note that we can choose different weights for averaging gyroscope and accelerometers.}

@@

When selecting weight for gyroscope or for accelerometer when the selection of vehicle frame is flexible, we can minimize the output noise or bias drift rate. Below we take the bias drift as an example but similar idea can be applied to the output noise.

The combined drift rate is as follow. Here we drop the subscript as it is the same for both gyroscope drift and accelerometer drift.
\begin{equation}
    \dot{\bar{\textbf{b}}}(t) = \sum_i{a_i \textbf{C}_i \textbf{w}_i(t)}
\end{equation}

The covariance of the combined drift rate would be the weighted sum of the noise covariance of each IMU.
\begin{equation}
    \bar{\bm{\Sigma}} = \sum_i{a_i^2 \bm{\Sigma}_i}
\end{equation}

@@

Since most IMU have isotropic noise for all three axes, the combined variance simplifies to,
\begin{equation}\label{noise_reduction_asym}
    \bar{\sigma}^2 = \sum_i{a_i^2 \sigma_i^2}
\end{equation}

Further more, if using $n$ same IMU and equation weighting, the combined variance is simply,
\begin{equation}\label{noise_reduction_sym}
    \bar{\sigma}^2 = \sum_i{a_i^2 \sigma^2} = n \left(\frac{\sigma}{n}\right)^2 = \frac{\sigma^2}{n}
\end{equation}

\subsection{Selecting Weight for Vehicle Frame Placement}

Given that $\sum_i{a_i \textbf{r}_i} = \textbf{0}$, if there are only two IMUs, we can only place the vehicle frame on the line through the center of the two IMUs. If there are three IMUs, we can only place the vehicle frame on the plane formed by the three IMUs. If there are four non-coplanar IMUs, then in theory, we can place the vehicle frame anywhere. However, placing the vehicle frame outside of the convex hull of the IMUs will increase the output noise which is usually not desired.

@@

Note that, in this section we will be using $w$ as weight instead of $a$ for convenience. Here $\textbf{r}_i$ is the relative position of the IMUs to the selected vehicle frame. Here we used the objective to reduce output noise as shown in section \ref{solve_weight_by_noise}. Since it is a quadratic programming problem, it is possible to find a close-form solution.

@@

Lastly since $\sum_i{w_i} = 1$ we need to normalize the intermediate solution to get the final weight for accelerometer.
\begin{equation}
    \textbf{w}^* = \frac{\hat{\textbf{w}}}{\textbf{1}^T \hat{\textbf{w}}}
\end{equation}

@@

@@

First we test our finding in simulation. To generate the simulated data, we first use combination of sinusoidal functions to generate the ground truth angular velocity and linear acceleration in body frame as shown in Figure \ref{fig:omega} and Figure \ref{fig:accel}. Then we integrate it to generate the ground truth trajectory.
% as shown in Figure \ref{fig:trajectory}
For each IMU, we simulate its measurement given the relative transformation in body frame and added noises and biases. Each IMU is sampled at 100 Hz. We selected MPU6050, a common MEMS IMU in the mechatronics community as the example model. All of the above is done using utilities like \texttt{kinematicTrajectory} and \texttt{imuSensor} provided by MATLAB.

We also simulate observations of known landmark from a monocular camera. This turns the estimation problem into a localization problem such that the error is bounded. For each camera frame, we first sample random observations in imaging plane as well as depth (between 2 and 10 m). We then perform reprojection and transform it to inertial frame with the ground truth pose to generate the landmark position. Camera observations arrives at 2 Hz. We selected such a low update rate to allow error to propagate to better show the benefits of averaging.

% \begin{figure*}[ht]
%     \centering
%     \begin{subfigure}[b]{0.4\linewidth}
%         \centering
%         \includegraphics[width=5.8cm]{figures/simulated_omega.png}
%         \caption{Simulated Angular Velocity}
%         \label{fig:omega}
%     \end{subfigure}
%     \begin{subfigure}[b]{0.4\linewidth}
%         \centering
%         \includegraphics[width=5.8cm]{figures/simulated_accel.png}
%         \caption{Simulated Linear Acceleration}
%         \label{fig:accel}
%     \end{subfigure}
%     \begin{subfigure}[b]{0.4\linewidth}
%         \centering
%         \includegraphics[width=5.8cm]{figures/simulated_trajectory.png}
%         \caption{Ground truth Trajectory from Integration}
%         \label{fig:trajectory}
%     \end{subfigure}
% \caption{Simulated Data Generation with MATLAB}
% \label{fig:switching_diagram}
% \end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=0.85\linewidth]{figures/simulated_omega.png}
    \caption{Simulated Angular Velocity}
    \label{fig:omega}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.85\linewidth]{figures/simulated_accel.png}
    \caption{Simulated Linear Acceleration}
    \label{fig:accel}
\end{figure}
% \begin{figure}
%     \centering
%     \includegraphics[width=0.85\linewidth]{figures/simulated_trajectory.png}
%     \caption{Ground truth Trajectory from Integration}
%     \label{fig:trajectory}
% \end{figure}

@@

@@

The results are summarized in Table \ref{tab:vio_result}. Example trajectories against ground truth are plotted in \textcolor{red}{Figure} \ref{}.

From the result, different configurations perform similarly. Error reduction from averaging as in simulation is not seen. This is believed to be due to various imperfections from the real-world dataset.

>>>>

% \title{Weighted Averaging of Multiple Non-collocated Inertial Measurement Units}
\title{Tunable VIMU frame from averaging multiple Non-collocated IMUs}

@@

\begin{abstract}
Multi-IMU fusion has been an effective way of improving MEMS IMU accuracy. In this article, we discuss an averaging technique for multiple IMUs with physical separation. We also present methods of selecting weighting terms such that we can minimize the uncertainty in fused sensor output and select the placement of the reference frame of the combined virtual IMU (VIMU). We tested our method in simulation and validated it on a real dataset. The results show that the averaging technique works for IMUs with large separation and performance gain is observed in simulation compared to using only a single IMU.
\end{abstract}

@@

An inertial measurement unit combining a gyroscope and accelerometer that provides measurement of angular velocity and linear acceleration has wide application in robotics and UAS. By integrating the measurement, one can track the relative orientation and position of the system to certain degree of accuracy within reasonable time frame \cite{tang2022_preintegration}. However due to the interoceptive nature of an IMU, it is often paired with additional sensors such as camera, lidar, or GNSS to reduce drift \cite{alaba2024_gps_and_imu, shan2020_liosam, he2020_visual_imu_and_gps}. In general, there are two main approaches to the tracking problem with an IMU: filter-based and graph optimization with pre-integrated IMU measurements.

In the filter-based approach, the IMU measurement is used as process input to an Extended Kalman Filter in the propagation step. Then the measurements from other exteroceptive sensors are used for state correction in the update step. One such example is the MSCKF (Multi-State Constraint Kalman Filter) proposed in \cite{Anastasios2007_MSCKF}, which uses the IMU for propagation and a camera for correction. It proposes a measurement model that directly relates two robot poses based on the landmark observation in the camera but independent from the landmark position itself. This allows for efficient tracking without including the positions of landmarks as part of the state vector as in traditional EKF-SLAM.

For the graph-optimization approach, such as VINS-MONO \cite{qin2018_vins-mono}, a pose graph is constructed by solving tightly coupled, nonlinear optimization problem with preintegrated IMU measurement and camera observations. Pose-graph optimization is run at the end to ensure global consistency.

Among many sensor configurations with an IMU, monocular camera plus IMU is one of the minimal configuration that is well-studied \cite{qin2018_vins-mono, 10616216}. Also IMU has been used heavily as a complementary sensor for GPS in GNSS-denied environment \cite{1008998, 8987949}. Based on these minimal configurations, various extension has been made. UMS-VINS \cite{jiang2023_UMS-VINS} provides a unified framework for using monocular and stereo cameras in visual-inertial odometry. Also the MSCKF framework \cite{Anastasios2007_MSCKF} works for multiple cameras with an IMU. There are also works that aim to provide a complete framework to incorporate multiple different sensors in one unified system \cite{10587194}.

To improve robustness, there have being works in introducing redundant IMUs for fault detection to build fault tolerance system. To detect a measurement error, \cite{Sturza1988_redundant} proposed parity vector based on hypothesis testing. Optimal geometric configuration of multiple redundant IMUs is also explored in \cite{Colomina2004REDUNDANTIF, guerrier2009, xue2023}.

Despite works mentioned above, most systems and applications still only use single IMU in the estimation pipeline. The vehicle frame is choose to be at the frame of reference of the IMU to avoid introducing additional kinematics terms from the lever arm effect.

In this article we present that through weighted averaging of multiple non-collocated IMUs, we can choose arbitrary VIMU frame of reference within solvable domain.

@@

In this section, we will introduce some prior works on fusing multiple IMUs for use in estimation pipeline.

@@

One category of approach is to perform fusion of multiple IMUs before using the measurement in inertial navigation system (INS). For example, \cite{jafari2014_PEM} uses Prediction Error Minimization method to model noise of IMUs and use Kalman filter to estimate the optimal process signal. \cite{xue2023} proposed to use Kalman filter to estimate the angular velocity measured by multiple non-orthogonal gyroscopes. Instead of using Kalman filter, an easier approach would be to directly average multiple gyroscopes after aligned under a common virtual IMU frame as discussed in \cite{waegli2008, patel2022_multi-imu, Colomina2004REDUNDANTIF}.

However, averaging of multiple accelerometers was not considered due to the additional acceleration terms from rotation (lever arm effect), unless they are closely placed and the separations in between are ignored.

@@

Typically state estimation problem with a IMU involves tracking the orientation, linear velocity and linear acceleration of the vehicle (commonly defined at the IMU). With additional IMUs, additional state variables like linear acceleration, angular velocity and angular acceleration are added. This allows using the full mechanization model of accelerometer as measurement update in Kalman filter \cite{Bancroft2011DataFA, Beaudoin2018_satelite}.

@@

Another approach is to track the state of each IMU individually and associate them with geometric constraints \cite{waegli2008, Beaudoin2018_satelite}. Using multiple IMUs and cameras, \cite{Eckenhoff2021_MIMC-VINS} proposed to track all stacked states of base and auxillary IMUs using MSCKF \cite{Anastasios2007_MSCKF}. Assuming the extrinsic calibrations of all IMUs are known and fixed, relative geometric constraints is imposed between states of different IMUs in the form of measurement update in addition to camera observations.

@@

As presented, multiple approaches have being proposed. \ref{constraint} would be computationally expensive if number of IMU are large as it has many states all in one Kalman filter. The averaging approach in \ref{VIMU} would be the simplest to implement and doesn't significantly alter the existing single IMU framework. However, due to the lever arm effect, averaging multiple accelerometers that are physically separated is not visited. In \ref{methodology} we present that it is still possible to average the accelerometers despite the separation and give a tunable VIMU frame as a result.

@@

In this discussion, we use the following four frames of reference. 1) Inertial frame $\bm{\mathcal{F}}_i$. 2) Body frame $\bm{\mathcal{F}}_b$ is for expressing sensor extrinsic calibration relative to some mechanical datum. 3) Sensor frame $\bm{\mathcal{F}}_s$ is the local frame of reference at each sensor. 4) Vehicle frame $\bm{\mathcal{F}}_v$ is for tracking vehicle pose in state estimation pipeline. In later section we use vehicle frame interchangeably with VIMU frame.

For any point $\textbf{p}$ or vector we define as $\textbf{r}_i^{pi}$ in inertial frame $\mathcal{F}_i$. For example, the vehicle position in inertial frame thus become $\textbf{r}_i^{vi}$.

When referring to the pose of a vehicle, it is the tuple of $\{\textbf{C}_{iv}, \textbf{r}_i^{vi}\}$ which forms the transformation matrix as follow,
$$
\textbf{T}_{iv} = \left[\begin{matrix}
    \textbf{C}_{iv} & \textbf{r}_i^{vi} \\
    \textbf{0} & 1
\end{matrix}\right]
$$

The pose can be used to transform point $\textbf{r}_v^{pv}$ expressed in vehicle frame $\bm{\mathcal{F}}_v$ into inertial frame $\bm{\mathcal{F}}_i$ as $\textbf{r}_i^{pi}$:
$$
\left[\begin{matrix} \textbf{r}_i^{pi} \\ 1 \end{matrix}\right]
    = \textbf{T}_{iv} \left[\begin{matrix} \textbf{r}_v^{pv} \\ 1 \end{matrix}\right]
\quad \text{or} \quad
\textbf{r}_i^{pi} = \textbf{C}_{iv} \textbf{r}_v^{pv} + \textbf{r}_i^{vi}
$$

In this document $*^\wedge$ is specifically used for expressing a $\mathbb{R}^3$ as skew symmetric matrix.
$$
\bm{\omega}^\wedge = \left[\begin{matrix}
    0 & -\omega_2 & \omega_1 \\
    \omega_2 & 0 & -\omega_3 \\
    -\omega_1 & \omega_3 & 0
\end{matrix}\right]
$$

If a quantity is in the form $x(t)$ then it is a time-varying quantity, otherwise it is time-invariant.

@@

In addition, $\textbf{g}$ is the acceleration from gravity and $\bm{\omega}$ and $\textbf{a}$ are the true angular velocity and true linear acceleration in vehicle frame.

@@

Where $\textbf{b}_\omega$ is the bias of gyroscope measurement due to random walk and $\textbf{n}_\omega$ is the additive white noise of the measurement. Similarly we have the following measurement model for accelerometer at vehicle frame. Here $\textbf{b}_a$ is the bias of accelerometer and $\textbf{n}_a$ is the noise of the acceleration measurement.

@@

For IMU not at vehicle frame, measurement are in each IMU's local frame and we have additional terms in acceleration measurement due to lever arm effect.
\begin{equation}
    \textbf{y}_{\omega,i}(t) = \textbf{C}_{i}^{-1} \bm{\omega}(t) + \textbf{b}_{\omega, i}(t) + \textbf{n}_{\omega,i}(t)
\end{equation}
\begin{equation}
\begin{split}
    \textbf{y}_{a,i}(t) &= \textbf{C}_i^{-1} \left[ \textbf{a}(t) + \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i) + \bm{\alpha}(t)^\wedge \textbf{r}_i \right] \\
    &+ \textbf{b}_{a,i}(t) + \textbf{n}_{a,i}(t)
\end{split}
\end{equation}

Here we have $\textbf{C}_i$ and $\textbf{r}_i$ short for $\textbf{C}_{vs}$ and $\textbf{r}_v^{sv}$ defining the pose $\textbf{T}_{vs}$ of the $i$-th IMU in vehicle frame.

@@

First we align the gyroscope measurement with the vehicle frame.
\begin{equation}
    \textbf{C}_{i} \textbf{y}_{\omega,i}(t) = \bm{\omega}(t) + \textbf{C}_{i} \textbf{b}_{\omega, i}(t) + \textbf{C}_{i} \textbf{n}_{\omega,i}(t)
\end{equation}

The weighted average ($\sum_i{w_i} = 1$) of all gyroscope is simply,
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_\omega(t) &= \sum_i{w_i \textbf{C}_{i} \textbf{y}_\omega(t)} \\
    &= \sum_i{w_i \left( \bm{\omega}(t) + \textbf{C}_{i} \textbf{b}_{\omega,i}(t) + \textbf{C}_{i} \textbf{n}_{\omega,i}(t) \right)} \\
    &= \bm{\omega}(t) + \sum_i{w_i \textbf{C}_{i} \textbf{b}_{\omega,i}(t)} + \sum_i{w_i \textbf{C}_{i} \textbf{n}_{\omega,i}(t)}
\end{split}
\end{equation}

@@

Align all accelerometer measurements to vehicle frame we have,
\begin{equation}\label{eqn_accel}
\begin{split}
    \textbf{C}_{i} \textbf{y}_{a,i}(t) &= \textbf{a}(t) + \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i) + \bm{\alpha}(t)^\wedge \textbf{r}_i \\
    &+ \textbf{C}_{i} \textbf{b}_{a,i}(t) + \textbf{C}_{i} \textbf{n}_{a,i}(t)
\end{split}
\end{equation}

Summing equation (\ref{eqn_accel}) of all accelerometers gives,
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_a(t) &= \sum_i{w_i \textbf{C}_{i} \textbf{y}_a(t)} \\
    &= \textbf{a}(t) + \sum_i{w_i \textbf{C}_{i} \textbf{b}_{a,i}(t)} + \sum_i{w_i \textbf{C}_{i} \textbf{n}_{a,i}(t)} \\
    &+ \sum_i{w_i \bm{\omega}(t)^\wedge (\bm{\omega}(t)^\wedge \textbf{r}_i)} + \sum_i{w_i \bm{\alpha}(t)^\wedge \textbf{r}_i} \\
\end{split}
\end{equation}

Using the associate rule $\textbf{a}^\wedge \textbf{b} + \textbf{a}^\wedge \textbf{c} = \textbf{a}^\wedge \left(\textbf{b} + \textbf{c}\right)$, we have,
\begin{equation}
\begin{split}
    \bar{\textbf{y}}_a(t) &= \sum_i{w_i \textbf{C}_{i} \textbf{y}_a(t)} \\
    &= \textbf{a}(t) + \sum_i{w_i \textbf{C}_{i} \textbf{b}_{a,i}(t)} + \sum_i{w_i \textbf{C}_{i} \textbf{n}_{a,i}(t)} \\
    &+ \bm{\omega}(t)^\wedge \left(\bm{\omega}(t)^\wedge \left( \sum_i{w_i \textbf{r}_i} \right) \right) + \bm{\alpha}(t)^\wedge \left( \sum_i{w_i\textbf{r}_i} \right)\\
\end{split}
\end{equation}

We see that if $\sum_i{w_i \textbf{r}_i} = \textbf{0}$ then we can eliminate the last two terms from lever arm effect. This gives,
\begin{equation}
    \bar{\textbf{y}}_a(t) = \textbf{a}(t) + \bar{\textbf{b}}_a(t) + \bar{\textbf{n}}_a(t)
\end{equation}

@@

As we can see, if we choose the vehicle frame and weights of accelerometers such that $\sum_i{w_i \textbf{r}_i} = \textbf{0}$, then we can ignore the separation of all IMUs and fuse the accelerometer measurement through averaging.

Since gyroscope measurement is location independent, we can choose different weights for averaging gyroscope and accelerometers.

@@

When selecting weight for gyroscopes, or for accelerometers under the condition that the selection of vehicle frame is flexible, we can minimize the output noise or bias drift rate. Below we take the bias drift as an example but similar idea can be applied to the output noise.

The combined drift rate is as follow. Here we drop the subscript as it is the same for both gyroscope drift and accelerometer drift.
\begin{equation}
    \dot{\bar{\textbf{b}}}(t) = \sum_i{w_i \textbf{C}_i \textbf{w}_i(t)}
\end{equation}

The covariance of the combined drift rate would be the weighted sum of the noise covariance of each IMU.
\begin{equation}
    \bar{\bm{\Sigma}} = \sum_i{w_i^2 \bm{\Sigma}_i}
\end{equation}

@@

Since most IMU have isotropic noise for all three axes, the combined variance simplifies to,
\begin{equation}\label{noise_reduction_asym}
    \bar{\sigma}^2 = \sum_i{w_i^2 \sigma_i^2}
\end{equation}

Further more, if using $n$ same IMU and equation weighting, the combined variance is simply,
\begin{equation}\label{noise_reduction_sym}
    \bar{\sigma}^2 = \sum_i{w_i^2 \sigma^2} = n \left(\frac{\sigma}{n}\right)^2 = \frac{\sigma^2}{n}
\end{equation}

\subsection{Selecting Weight for VIMU Frame Placement}

Given that $\sum_i{w_i \textbf{r}_i} = \textbf{0}$, if there are only two IMUs, we can only place the vehicle frame on the line through the center of the two IMUs. If there are three IMUs, we can only place the vehicle frame on the plane formed by the three IMUs. If there are four non-coplanar IMUs, then in theory, we can place the vehicle frame anywhere. However, placing the vehicle frame outside of the convex hull of the IMUs will increase the output noise which is usually not desired.

@@

Here $\textbf{r}_i$ is the relative position of the IMUs to the selected vehicle frame. We used the objective to reduce output noise as shown in section \ref{solve_weight_by_noise}. Since it is a quadratic programming problem, it is possible to find a close-form solution.

@@

Lastly since $\sum_i{w_i} = 1$ we need to normalize the intermediate solution to get the final weights for accelerometers.
\begin{equation}
    \textbf{w}^* = \frac{\hat{\textbf{w}}}{\textbf{1}^T \hat{\textbf{w}}}
\end{equation}

@@

To check the correctness of the frame selection method through weighted averaging and to show performance benefit of the fused VIMU we first performed evaluation in simulation.

@@

To generate the simulated data, we first use combination of sinusoidal functions to generate the ground truth angular velocity and linear acceleration in body frame as shown in Figure \ref{fig:omega} and Figure \ref{fig:accel}. Then we integrate them to generate the ground truth trajectory. We selected MPU6050, a common MEMS IMU in the mechatronics community as the example model. For each IMU, we simulate its measurements given the relative transformation in body frame and added noises and biases. Each IMU is sampled at 100 Hz. All of the above is done using utilities like \texttt{kinematicTrajectory} and \texttt{imuSensor} provided by MATLAB.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{figures/simulated_omega.png}
    \caption{Simulated Angular Velocity}
    \label{fig:omega}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{figures/simulated_accel.png}
    \caption{Simulated Linear Acceleration}
    \label{fig:accel}
\end{figure}
% \begin{figure}
%     \centering
%     \includegraphics[width=0.85\linewidth]{figures/simulated_trajectory.png}
%     \caption{Ground truth Trajectory from Integration}
%     \label{fig:trajectory}
% \end{figure}

@@

We also simulate observations of known landmark from a monocular camera. This turns the estimation problem into a localization problem such that the error is bounded. For each camera frame, we first sample random observations in imaging plane as well as depth (between 2 and 10 m). We then perform reprojection and transform it to inertial frame with the ground truth pose to generate the corresponding landmark positions. Camera observations arrives at 2 Hz. We selected this low update rate to allow error to propagate for better showing the benefits of averaging.

@@

The results are summarized in Table \ref{tab:vio_result}. Example trajectories against ground truth are plotted in Figure \ref{fig:posyaw_af}.
